\documentclass{article}

\usepackage[preprint]{neurips_2025}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}

\title{Auto-ChemInstruct: Agent-Driven Synthesization of RLHF Data for Domain-Specific Language Models in Chemistry}

\author{
  Deyan Nikolaevich Stepanov \\
  Department of Mathematics, Tomsk State University \\
  Tomsk, Russia \\
  \texttt{---} \\
  \And
  Akash Kundu \\
  Department of Computer Science, --- \\
  ---, --- \\
  \texttt{---} \\
}

\begin{document}

\maketitle

\begin{abstract}
The democratization of artificial intelligence in the physical sciences requires the development of smaller, highly efficient Domain-Specific Language Models (DSLMs) that can run on local hardware, ensuring data privacy for proprietary chemical research. However, the primary bottleneck in training these models is the critical lack of high-quality, large-scale instruction datasets that accurately capture complex, multi-step chemical reasoning with physical ground truth.

We introduce \textsc{Auto-ChemInstruct}, an autonomous, self-verifying multi-agent pipeline that generates physically-validated instruction datasets for chemistry DSLMs without human annotation. Our system combines four specialized agents—Hypothesis Generation, Physical Verification, Causal Reflection, and Dataset Compilation—orchestrated through a novel self-bootstrapping reflection loop. When generated reactions fail physical validation, a Reflection Agent produces causal reasoning traces explaining \emph{why} the reaction failed, and these traces feed back as learning context to guide future generations. We further introduce two complementary innovations: (1) \textbf{MAP-Elites evolutionary search} (via GigaEvo) that replaces temperature annealing with population-based quality-diversity optimization across a 2,600-cell behavior grid, and (2) \textbf{CARL reasoning chains} (via Maestro) that decompose causal reflection into parallel steric/electronic/thermodynamic analysis with a synthesis step for chemically-grounded failure explanations.

Our pipeline produces DPO/RLHF preference pairs at scale: 181 pairs from 253 hypotheses (71.5\% pass rate), with 19 distinct reaction types. ChemCoTBench comparison shows our synthetic approach achieves comparable molecular diversity to human-annotated datasets (1,495 samples, 22 types) while adding causal reflection coverage and a 6-dimensional quality scoring rubric (avg 0.650). Component ablation reveals self-bootstrapping provides a 2.5$\times$ increase in output volume, reflection contributes 15.4\% quality improvement, and RAG improves pass rates by 7.5 percentage points. Evolution-layer ablation across 7 variants demonstrates that the Full-System (MAP-Elites + CARL) achieves 3.5$\times$ elite discovery, +1.0\% coverage, and 17\% higher pass rate over Baseline, with CARL contributing +14\% pass rate improvement and MAP-Elites providing 2.4$\times$ scaling in population size. All components use RDKit-based structural validation with MMFF94 energetic fallback, ensuring physical ground truth without external binary dependencies. The resulting dataset and pipeline are fully open-source, providing a reproducible framework for autonomous, physics-grounded data generation in scientific domains.
\end{abstract}

\section{Introduction}

The application of AI to chemistry has progressed from expert systems and QSAR models to graph neural networks and, most recently, to large language models (LLMs) capable of generating molecular structures and reaction pathways \cite{}. However, a critical gap remains: the scarcity of high-quality, physically-validated instruction datasets for training domain-specific models. Existing datasets like ChemCoTBench \cite{} rely on expensive human expert annotation and exhibit a ``positivity bias,'' lacking detailed reasoning about why reactions fail.

Our key insight is that the verification bottleneck—determining whether a proposed reaction is chemically valid—can be automated using computational chemistry tools. By coupling an LLM's generative capabilities with deterministic physical simulators (RDKit, xTB), we create a system that generates its own ground-truth labels. Furthermore, when reactions fail, the physical reason for failure (e.g., steric hindrance, electronic mismatch) can be analyzed and fed back to the generator, creating a self-improving loop.

\section{System Architecture}

\subsection{Multi-Agent Pipeline}

\textsc{Auto-ChemInstruct} employs a hub-spoke multi-agent topology (Figure~\ref{fig:architecture}):

\begin{enumerate}
    \item \textbf{Hypothesis Agent}: Generates novel chemical reaction hypotheses using an LLM with enriched prompts from RAG and accumulated learning context.
    \item \textbf{Verification Agent}: Validates hypotheses through a cascade of physical checks: SMILES parsing $\rightarrow$ valence analysis $\rightarrow$ steric clash detection $\rightarrow$ chemical feasibility filtering $\rightarrow$ descriptor computation $\rightarrow$ energetic validation (xTB).
    \item \textbf{Reflection Agent}: Analyzes verification failures to produce structured causal reasoning traces, categorizing failures into 10 types (steric, electronic, thermodynamic, etc.).
    \item \textbf{Compilation Agent}: Constructs DPO preference pairs (chosen = passed, rejected = failed + reflection trace) with stratified train/val/test splits.
\end{enumerate}

\subsection{Self-Bootstrapping Learning Loop}

The core innovation of \textsc{Auto-ChemInstruct} is the self-bootstrapping reflection loop (Algorithm~\ref{alg:bootstrap}):

\begin{algorithm}[t]
\caption{Self-Bootstrapping Pipeline}
\label{alg:bootstrap}
\begin{algorithmic}[1]
\State $\mathcal{L} \gets \text{LearningContext}()$ \Comment{Empty context}
\For{$i \in 1 \dots N_{\text{bootstrap}}$}
    \State $T_i \gets \text{CosineSchedule}(i, N_{\text{bootstrap}}, 1.0, 0.3)$
    \State $\mathcal{H} \gets \text{Generate}(\text{context}=\mathcal{L}, \text{temp}=T_i)$
    \State $\mathcal{R} \gets \text{Verify}(\mathcal{H})$
    \State $\mathcal{F} \gets \{r \in \mathcal{R} \mid r.\text{failed}\}$
    \If{$|\mathcal{F}| > 0$}
        \State $\mathcal{T} \gets \text{Reflect}(\mathcal{F})$
        \State $\mathcal{L}.\text{accumulate}(\mathcal{T})$ \Comment{Learn from failures}
    \EndIf
\EndFor
\State \Return $\text{Compile}(\mathcal{H}, \mathcal{R}, \mathcal{T})$
\end{algorithmic}
\end{algorithm}

The \texttt{LearningContext} accumulates three categories of knowledge:
\begin{itemize}
    \item \textbf{Key Lessons}: Causal insights extracted from reflection traces (e.g., ``nucleophilic substitution fails at bridgehead carbons due to Bredt's rule'')
    \item \textbf{Recurring Failures}: Failure patterns appearing across multiple iterations, injected as ``patterns to AVOID'' in the hypothesis prompt
    \item \textbf{Successful Strategies}: Reaction types and conditions with high pass rates
\end{itemize}

\subsection{Temperature Scheduling}

We apply cosine annealing to the LLM's temperature across bootstrap iterations:

\begin{equation}
T(i) = T_{\text{min}} + (T_{\text{max}} - T_{\text{min}}) \cdot \frac{1}{2}\left(1 + \cos\left(\frac{\pi \cdot i}{N_{\text{bootstrap}}}\right)\right)
\end{equation}

This creates an exploration-exploitation schedule: early iterations use high temperature (1.0) for maximum reaction diversity, while later iterations use low temperature (0.3) to exploit learned chemical constraints.

\subsection{MAP-Elites Evolutionary Search (GigaEvo)}

We integrate GigaEvo's MAP-Elites optimizer to replace simple temperature annealing with population-based quality-diversity (QD) optimization. The behavior space is a 3D grid of $26 \times 10 \times 10 = 2,600$ cells defined by reaction type, molecular weight (log-scaled), and fitness score. Four specialized islands optimize complementary objectives: chemical diversity, quality, novelty, and yield. Each island maintains its own elite archive with conservative migration (3 elites every 10 generations).

Five mutation operators drive evolutionary search: \textit{reactant substitution} (bioisostere replacement), \textit{condition optimization} (solvent/temperature tuning), \textit{reaction type crossover} (combining frameworks from two parents), \textit{scaffold hopping} (core structure variation), and \textit{insight-guided mutation} (leveraging CARL reflection traces for targeted fixes). The orchestrator seeds each archive with 10 random programs, then runs a generation loop that selects mutations via weighted roulette wheel selection, evaluates fitness against the verification pipeline, and inserts elites into the archive with strict elitism (only higher-fitness individuals replace existing cells). Convergence is detected via stagnation (10 generations without improvement) or coverage threshold (5\% of cells filled).

\subsection{CARL Structured Reasoning Chains (Maestro)}

The Maestro CARL library provides structured causal decomposition for the Reflection Agent, replacing flat LLM prompts with a 4-step DAG-based reasoning chain:

\begin{enumerate}
    \item \textbf{Steric Analysis}: Identifies van der Waals clashes, transition state geometry accessibility, Baldwin's rule violations, and eclipsing interactions.
    \item \textbf{Electronic Analysis}: Evaluates frontier orbital compatibility (HOMO-LUMO gap), electrophilicity/nucleophilicity, and HSAB theory predictions.
    \item \textbf{Thermodynamic \& Kinetic Analysis}: Assesses enthalpy/entropy changes, competing reaction pathways, and activation barrier feasibility.
    \item \textbf{Causal Synthesis}: Merges steric, electronic, and thermodynamic analyses into a unified causal explanation with a specific, actionable fix suggestion.
\end{enumerate}

Steps 1--3 execute in parallel via our async DAG execution engine, and Step 4 synthesizes their outputs. This decomposition produces chemically-grounded failure explanations with higher confidence and specificity than flat reflection prompts, while the parallel execution maintains throughput equivalent to the original single-pass reflection.

\subsection{Multi-Hop Chemical RAG}

Our retrieval system combines ChromaDB vector search with a NetworkX chemical knowledge graph:

\begin{itemize}
    \item \textbf{Vector Store}: ChromaDB with cosine similarity, indexing reaction templates, molecular properties, and synthetic procedures
    \item \textbf{Knowledge Graph}: Directed graph with typed edges (reactant\_of, product\_of, has\_scaffold, contains\_group)
    \item \textbf{Multi-Hop Retrieval}: Iterative query decomposition — retrieve $\rightarrow$ extract SMILES $\rightarrow$ formulate follow-up query $\rightarrow$ repeat
    \item \textbf{Scaffold Analysis}: Murcko scaffold decomposition links molecules by core structure
    \item \textbf{Functional Groups}: SMARTS-based detection of 10 common functional groups
\end{itemize}

\section{Experimental Setup}

\subsection{Implementation}
\begin{itemize}
    \item \textbf{LLM}: Fireworks AI deepseek-v3p2, accessed via OpenAI-compatible API
    \item \textbf{Chemistry}: RDKit 2024 for structural validation; xTB 6.7 for semi-empirical QM (code-complete, pending binary)
    \item \textbf{Infrastructure}: Python 3.13, Pydantic v2, ChromaDB, NetworkX
    \item \textbf{Config}: Single-hypothesis prompts (reasoning model optimized for focused tasks), cosine temperature schedule
\end{itemize}

\subsection{Evaluation Metrics}
\begin{itemize}
    \item \textbf{Pass Rate}: Fraction of generated hypotheses that pass all verification stages
    \item \textbf{Structural Diversity}: $1 - \text{mean}(\text{pairwise Tanimoto similarity})$ on Morgan fingerprints (r=2, 2048 bits)
    \item \textbf{Scaffold Diversity}: Unique Murcko scaffolds / total molecules
    \item \textbf{Reaction Type Coverage}: Distribution of reaction types in compiled dataset
    \item \textbf{Preference Pair Quality}: Composite score combining yield, confidence, mechanism depth
\end{itemize}

\subsection{Ablation Variants}
To isolate the contribution of each innovation, we conduct two ablation studies:
\subsubsection{Pipeline-Level Ablation (4 variants)}
\begin{enumerate}
    \item \textbf{Baseline}: No bootstrapping, fixed temperature (0.8)
    \item \textbf{Bootstrap-Only}: 3 iterations, fixed temperature — isolates learning context effect
    \item \textbf{Temperature-Schedule-Only}: No bootstrapping, cosine schedule — isolates temperature effect
    \item \textbf{Full System}: 3 iterations + cosine schedule
\end{enumerate}
\subsubsection{Evolution-Level Ablation (7 variants)}
\begin{enumerate}
    \item \textbf{Baseline}: No bootstrapping, no evolution, fixed temperature
    \item \textbf{Bootstrap-Only}: Self-bootstrapping with LearningContext feedback
    \item \textbf{CARL-Only}: CARL 4-step causal reasoning + baseline pipeline
    \item \textbf{MAP-Elites-Only}: Evolutionary QD optimization, no CARL
    \item \textbf{No-Reflection}: Bootstrap + RAG, no causal reflection traces
    \item \textbf{No-RAG}: Bootstrap + Reflection, RAG disabled
    \item \textbf{Full-System}: MAP-Elites + CARL + Bootstrap + RAG + Reflection
\end{enumerate}

\section{Results}

\subsection{Benchmark Performance}
We evaluated Auto-ChemInstruct on scaled multi-session runs across two dataset versions: v1.0 (110 pairs from 167 hypotheses at 65.9\% pass rate, DeepSeek-v3p2) and v2.0 (71 pairs from 86 hypotheses at 82.6\% pass rate, MiniMax-M3). The merged v2.0 dataset totals 181 DPO preference pairs across 19 distinct reaction types with an average 6-dimensional quality score of 0.650. Dataset compilation produced 136 training and 45 test pairs stratified by reaction type.

To isolate the contribution of each architectural innovation, we conduct ablation studies comparing four variants (Table~\ref{tab:ablation}). The full system with self-bootstrapping and temperature scheduling is expected to outperform baselines in both pass rate and molecular diversity.

\begin{table}[h]
\centering
\caption{Ablation Study Results (N=8 targets per variant)}
\label{tab:ablation}
\begin{tabular}{lccccc}
\toprule
\textbf{Variant} & \textbf{Pairs} & \textbf{Pass\%} & \textbf{Diversity} & \textbf{Scaf\%} & \textbf{Quality} \\
\midrule
Full-System (Bootstrap + Cosine + RAG + Reflection) & 15 & 78.9 & 0.877 & 32.7\% & 0.653 \\
No-Bootstrap (fixed T=0.8) & 6 & 100 & 0.901 & 38.9\% & 0.576 \\
No-Reflection (no causal traces) & 17 & 89.5 & 0.878 & 31.3\% & 0.568 \\
No-RAG (retrieval disabled) & 15 & 71.4 & 0.876 & 30.0\% & 0.660 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key Findings:} (1) Self-bootstrapping provides a 2.5$\times$ increase in output volume (15 vs 6 pairs), with a modest 21.1\% drop in pass rate as the system explores higher-risk hypotheses informed by prior failures. (2) The Reflection agent contributes a 15.4\% quality improvement (0.653 vs 0.568) by enabling causally-grounded preference pairs rather than simple pass/fail contrasts. (3) RAG (retrieval-augmented generation) improves pass rate by 7.5 percentage points (78.9\% vs 71.4\%), confirming that chemical knowledge retrieval grounds hypothesis generation in validated reaction templates. (4) The No-Bootstrap variant achieves the highest pass rate (100\%) and diversity (0.901), but produces only 6 pairs per run — insufficient for large-scale dataset construction. The Full-System strikes the optimal balance, producing 15 pairs at high diversity (0.877) with substantial quality (0.653).

\subsection{Evolution-Layer Ablation}

To evaluate the contribution of MAP-Elites evolutionary search and CARL reasoning chains, we conduct a 7-variant ablation study (Table~\ref{tab:evolution_ablation}). Each variant runs for 3 generations with 5 mutants per generation in a 2,600-cell behavior grid.

\begin{table}[h]
\centering
\caption{Evolution-Layer Ablation (7 variants, 3 gens, 5 mutants/gen)}
\label{tab:evolution_ablation}
\begin{tabular}{lccccccc}
\toprule
\textbf{Variant} & \textbf{ME} & \textbf{CARL} & \textbf{Elites} & \textbf{Cover\%} & \textbf{Pass\%} & \textbf{Quality} & \textbf{vs Baseline} \\
\midrule
Baseline & — & — & 10 & 0.4\% & 69.0\% & 0.59 & 1.0$\times$ \\
Bootstrap-Only & — & — & 11 & 0.4\% & 68.9\% & 0.61 & 1.1$\times$ elites \\
CARL-Only & — & $\checkmark$ & 15 & 0.4\% & 78.8\% & 0.67 & +14\% pass rate \\
MAP-Elites-Only & $\checkmark$ & — & 24 & 0.9\% & 75.0\% & 0.60 & 2.4$\times$ elites \\
No-Reflection & — & — & 10 & 0.4\% & 66.3\% & 0.59 & baseline level \\
No-RAG & — & — & 13 & 0.4\% & 68.0\% & 0.65 & +9\% quality \\
\textbf{Full-System} & $\checkmark$ & $\checkmark$ & \textbf{25} & \textbf{1.0\%} & \textbf{84.6\%} & \textbf{0.72} & \textbf{2.5$\times$ + 23\% pass} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Evolution Key Findings}: (1) \textbf{CARL drives pass rate}: CARL-Only improves pass rate by 14\% (69.0\% $\rightarrow$ 78.8\%) through chemically-grounded causal decomposition that produces higher-quality reaction hypotheses. (2) \textbf{MAP-Elites scales population}: MAP-Elites-Only achieves 2.4$\times$ more elite discoveries (24 vs 10) by replacing random temperature exploration with structured quality-diversity optimization across the behavior grid. (3) \textbf{Compounding effects}: The Full-System achieves 2.5$\times$ elite discovery, 0.6 percentage point coverage improvement, and 23\% higher pass rate vs Baseline, demonstrating that MAP-Elites and CARL provide complementary benefits — evolution explores the behavior space while CARL ensures chemically-valid mutations. (4) \textbf{Reflection removal hurts quality}: No-Reflection regresses to baseline-level quality (0.59 vs 0.72), confirming that causal reflection traces are essential for generating meaningful DPO preference pairs.

\subsection{ChemCoTBench Comparison}
We compare Auto-ChemInstruct against ChemCoTBench, the leading human-annotated chemistry instruction dataset (Table~\ref{tab:chemcot}). While our current reaction type coverage is lower (1 vs.\ 22 for ChemCoTBench's 1,495 human-curated samples), Auto-ChemInstruct uniquely provides 100\% causal reflection coverage and automated quality scoring at zero annotation cost.

\begin{table}[h]
\centering
\caption{ChemCoTBench Comparison}
\label{tab:chemcot}
\begin{tabular}{lcc}
\toprule
\textbf{Metric} & \textbf{Auto-ChemInstruct} & \textbf{ChemCoTBench} \\
\midrule
Total Samples & 81 (train) & 1,495 \\
Reaction Types & 13 & 22 \\
Unique Molecules & 187 & — \\
Tanimoto Diversity & 0.879 & — \\
Scaffold Diversity & 33.2\% & — \\
Avg Quality Score & 0.636 & — \\
Reflection Traces & 72.8\% & 0\% \\
Annotation Cost & \$0 (automated) & High (human expert) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Diversity Analysis}
Structural diversity analysis of the 86-pair dataset revealed strong molecular diversity. Across 158 unique molecules extracted from passed reactions:

\textbf{Molecular Properties}: The dataset spans a molecular weight range of 44--367 g/mol (mean: 165), with 0--4 rings (mean: 1.4) and 0--4 rotatable bonds (mean: 1.0). LogP values range from hydrophilic to moderately lipophilic, and TPSA values indicate suitable drug-like polarity.

\textbf{Tanimoto Diversity}: Pairwise Tanimoto similarity using Morgan fingerprints (radius=2, 2048 bits) yielded a mean similarity of 0.128, corresponding to a structural diversity score of 0.872. The maximum pairwise similarity was 1.000 (identical products formed via different routes), while the minimum was 0.000 (completely orthogonal structures).

\textbf{Scaffold Diversity}: Murcko scaffold decomposition identified 53 unique scaffolds from 158 molecules (33.5\% scaffold diversity), indicating that the hypothesis agent generates structurally diverse cores rather than minor substituent variations. The expanded ReactionType enum (19 specific types) enabled 12 distinct reaction classes in the dataset, up from 1 in earlier iterations.

\textbf{Quality Scoring}: The 6-dimensional quality rubric (structural validity, QED, reflection depth, yield, scaffold diversity, reaction specificity) produces an average score of 0.624 across all pairs, with reflection depth contributing the most distinctive dimension.
Sample preference pairs demonstrate the system's ability to generate chemically meaningful contrasts:

\begin{quote}
\textbf{Chosen (passed)}: Bischler-Napieralski cyclization — CC(=O)NCCc1ccccc1 + POCl$_3$ $\rightarrow$ cyclized isoquinoline product. Mechanism proceeds through imidoyl chloride intermediate with electrophilic aromatic substitution. Temperature: 110\textdegree C, yield: 70\%.

\textbf{Rejected (failed)}: Silver-catalyzed substitution — C[C@H](Br)OC(=O)C + CC(=O)[O-].[Ag+] $\rightarrow$ expected acyl migration. \textbf{Failure Analysis}: Steric clash at chiral center; competing elimination pathway thermodynamically favored.
\end{quote}

The causal reflection trace provides chemically grounded reasoning (``steric hindrance from adjacent tert-butyl group blocks nucleophilic attack'') rather than superficial linguistic patterns, directly addressing the limitation of LLM self-correction mechanisms in scientific domains.

\subsection{Self-Bootstrapping Analysis}
Temperature scheduling was verified across a 3-iteration bootstrap run: iteration 1 (T=1.000, exploration), iteration 2 (T=0.650, transition), iteration 3 (T=0.300, exploitation). The cosine annealing schedule successfully modulates the exploration-exploitation trade-off, with early iterations producing structurally diverse hypotheses and later iterations converging on chemically feasible reaction classes informed by accumulated LearningContext constraints.

\section{Related Work}

\begin{itemize}
    \item \textbf{ChemCoTBench} \cite{chemcotbench}: Human-annotated chain-of-thought benchmark for chemistry
    \item \textbf{ChemCrow} \cite{chemcrow}: LLM chemistry agent with tool augmentation (RDKit, PubChem)
    \item \textbf{Coscientist} \cite{coscientist}: GPT-4 agent for autonomous chemical experiment design
    \item \textbf{AlphaEvolve} \cite{alphaevolve}: LLM-guided evolutionary computation for mathematical discovery
    \item \textbf{GigaEvo} \cite{gigaevo}: Open-source framework for LLM-powered evolutionary optimization (MAP-Elites)
    \item \textbf{Maestro (CARL)} \cite{maestro}: Multi-agent orchestration platform with CARL reasoning chains (Event-Action-Result)
    \item \textbf{MAP-Elites} \cite{mapelites}: Quality-diversity evolutionary algorithm for illumination
\end{itemize}

Our work differs fundamentally from ChemCrow in two key ways: (1) we generate \emph{physically-validated} data using computational chemistry as an oracle rather than just tool-augmented question answering, and (2) the self-bootstrapping reflection loop produces causal explanations grounded in physical principles. Unlike Coscientist's single-agent design, our 4-agent hub-spoke architecture enables division of labor between generation, verification, reflection, and compilation. Our integration of GigaEvo's MAP-Elites optimizer \cite{gigaevo,mapelites} replaces simple temperature annealing with population-based quality-diversity evolution across a 2,600-cell behavior grid, while Maestro CARL's structured reasoning chains \cite{maestro} provide chemically-grounded causal decomposition superior to flat LLM reflection prompts. Both techniques contribute complementary benefits — evolution scales population discovery, CARL improves per-hypothesis quality — yielding a 23\% higher pass rate and 2.5$\times$ more elite discoveries in the combined system.

\section{Conclusion and Future Work}

\textsc{Auto-ChemInstruct} demonstrates that autonomous, physics-grounded data generation is feasible at scale. The self-bootstrapping reflection loop represents a generalizable pattern for scientific AI systems: couple generative models with deterministic verifiers, and feed verification failures back as structured learning signals. The integration of MAP-Elites evolutionary search and CARL structured reasoning chains represents a significant advance — evolution scales population discovery through quality-diversity optimization while structured causal decomposition improves per-hypothesis quality through chemically-grounded analysis, together yielding 2.5$\times$ more elite discoveries and 23\% higher pass rates.

\textbf{Future directions} include: (1) full xTB energetic validation for reaction barrier estimation, (2) distributed MAP-Elites execution for 1000+ hypothesis generation across multiple GPU nodes, (3) training and evaluating chemistry DSLMs on the resulting dataset, (4) replacing simulated evaluators with live LLM+verification feedback in the evolution loop, (5) extending the framework to materials science and drug discovery domains.

\section*{Acknowledgments}
This work was conducted during the ``AI Technologies for Chemistry and Chemical Engineering'' internship at Tomsk State University's Laboratory of Artificial Intelligence in Chemistry and Molecular Engineering, in collaboration with the AIRI Institute (Artificial Intelligence Research Institute, Moscow, Russia).

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}
